---
title: 'Randomization Inference and Multiple Hypothesis Testing'
output:
  html_document:
    df_print: paged
fig_caption: yes
---

##  Loading pacakges
### key: ri2 is randomization inference package

```{r message = FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(stargazer, quietly=TRUE) # for creating tables
library(ggplot2, quietly = TRUE) # for creating figures
library(ri2, quietly = TRUE)  # package for randomization inference.
```

##  Setup, This is just doing the data preparation from Lab 1

```{r}
df <- read.table("/Users/ev6/Documents/GitHub/edward-vytlacil.github.io/Data/PROGRESA.csv", header=TRUE, sep=",") # import csv

df$treat <- df$progresa1 
df$girl <- df$sex1 

df<-subset(df, select= c(sooloca,sooind_id,age1,hgc1,school,treat,girl,wave))
df<-subset(df,age1>=6 & age1<=19 & (wave==5 | wave==2) )
df$age1 <- ifelse(df$age1>=6 & df$age1<=18, df$age1, NA)
df$post <- ifelse(df$wave==5, 1, 0)  
dfPre <- df[df$post==0,]  
dfPost <- df[df$post==1,]  
```

##  Defining Function to run randomization inference
### using conduct_ri and declare_ra functions
### from ri2 package
```{r }

no.reps <-3000  # defining constant, number of simulations to do (number of replications)
 
# 
# defining function for randomization inference
# argument is the data frame, make it easier to do 
# on alternative subsets of data.  This is different
# from your PS3 will do inference for alterantive 
# outcome variables.  
#

f.ri <- function(d){

# number of villages in data frame
n.0<- length(unique(d$sooloca)) 
# number of treated villages in data frame
m.0<- length(unique(subset(d,d$treat==1)$sooloca))

# declaring randomization 
# declaring randomization procedure to be at village level 
# that is why clusters set to village id
# and that they randomized in n.0 villages out of 
# m.0.
declaration_ra<-declare_ra(N=n.0,m=m.0,clusters=d$sooloca) 

# now conducting randomization inference using our declared randomization procedure.

ri_out <-  conduct_ri(
     test_function = 
      function(d){with(d, abs(mean(d[treat==1,]$school) - mean(d[treat==0,]$school)))},  #defining test statistic as absolute value of mean difference of outcome among treated vs. control
    declaration = declaration_ra, # declaring randomoization protocol as defined above.
    assignment = "treat", # treatment variable.
      outcome = "school", # outcome variable
    data = d, # data frame
    sims = no.reps # number of simulations/replications
  )
ri_out}
```

##  Conduction randomization inference
### Inference on different samples  (full sample, sample of boys, etc.)

```{r results='asis'}
# run randomization inference on full sample
# so function at data frame dfPost
ri.fullsample<- f.ri(dfPost)
summary(ri.fullsample)
plot(ri.fullsample)


# same as above, but now
# separately by boys vs girls
ri.bysex<- by(dfPost, dfPost$girl,f.ri)  # "by"" similar to tapply, applying f.ri to dfPost by levels of dfPost$girl
summary(ri.bysex[[1]]) # boys
summary(ri.bysex[[2]]) # girls
plot(ri.bysex[[1]]) +ggtitle("Randomization Inference of Sharp Null of No Effect, Boys")
plot(ri.bysex[[2]]) +ggtitle("Randomization Inference of Sharp Null of No Effect, Girls")

# same as above, but now
# separately by (sex,highest grade completed)

ri.bysexgrade <- by(dfPost, list(dfPost$girl,dfPost$hgc1),f.ri)
pvalues.boys <- sapply(1:10,function(j){summary(ri.bysexgrade[[1,j]])$two_tailed_p_value} ) 
pvalues.girls <- sapply(1:10,function(j){summary(ri.bysexgrade[[2,j]])$two_tailed_p_value} )
results.p<-cbind(1:10,pvalues.boys,pvalues.girls)
results.names <- c("Grade","p-value boys","p-value girls")
colnames(results.p)<-results.names

# Use stargazer to report a  table. 
stargazer(results.p,  digits=3, type="html", title="randomization inference p-values, inference by sex and grade")
 
#example plot, girls 6th grade
plot(ri.bysexgrade[[2,6]]) +ggtitle("Randomization Inference of Sharp Null of No Effect, Girls, Grade 6")
```


## Multiple hypothesis testing corrections
```{r, results='asis'}

varlabels  <- c(1:10)
 
f.adjust <- function(x,label){
  results.p<-cbind(x,p.adjust(x, method=c("bonferroni")),p.adjust(x, method=c("holm")))
  colnames(results.p)<-c("No Correction","Bonferroni","Holm's")
  rownames(results.p)<-label
   results.p
   }

stargazer(f.adjust(pvalues.boys,varlabels), type = "html", digits=3, title="P-Values: Boys, Testing Sharp Null of No Treatment Effect")

stargazer(f.adjust(pvalues.girls,varlabels), type = "html", digits=3, title="P-Values: Girls, Testing Sharp Null of No Treatment Effect")

```
